---
title: "Transmission dynamics in the era of test and treat" 
author: "Adam Akullian and Josh Herbeck"
date: "`r Sys.Date()`"
output: 
  html_document:
    df_print: paged
    highlight: tango
    toc: true
    toc_float: true
    theme: united
    fig_width: 7
    fig_height: 4
    fig_caption: true
    code_folding: hide
---
## Project description
We used output from a stochastic agent-based epidemic model calibrated to the HIV epidemic in the Kingdom of Eswatini (formerly Swaziland) to estimate the number and relative frequency of secondary infections caused by HIV infected individuals according to age group, sex, stage of infection (acute, early, latent, AIDS), and ART status. We estimated changes in the contribution of these sources to the epidemic over the previous and next thirty years, under multiple ART-defined scenarios (e.g. pre-ART, ART, ART scale up, status-quo ART coverage (current 90-90-90 targets), and increasing levels of UTT near-perfect ART initiation
Documentation on all model parameters: https://idmod.org/docs/hiv/software-report-transmission.html?searchText=TransmissionReport

# Setup
```{r setup, message = FALSE}
require("knitr")
knitr::opts_chunk$set(echo = TRUE, fig.align='center')
opts_knit$set(root.dir = "C:\\Users\\aakullian\\Dropbox (IDM)\\GitHub\\EMOD_eswatini\\Output\\")

library(kableExtra)
library(tidyverse)
library(janitor)
library(lubridate)
library(MASS)
library(glmnet)
library(parallel)
library(data.table)
library(networkD3)
library(RColorBrewer)
library(MESS)
numCores <- detectCores()

## Load helper functions
#source("../Scripts/model_helper_functions.R")

## Load output from EMOD
setwd("C:\\Users\\aakullian\\Dropbox (IDM)\\GitHub\\EMOD_eswatini\\Output")
load(file = "inputfiles2.rda") #brings in transmissionreport.csv and reporthivbyageandgender.csv

#create "date of acquisition" variable as date when the dest_ID acquired HIV (whether or not there was a subsequent transmission)
date.of.acquisition <- data.frame("SRC_ID"=transmissionreport$DEST_ID, "acq.year"=transmissionreport$YEAR, "sim.id"=transmissionreport$sim.id, "scenario"=transmissionreport$scenario) #source id used to merge in with source of transmitters
network.data <- left_join(transmissionreport, date.of.acquisition, by=c("SRC_ID","sim.id","scenario")) #join to include date of acquisition

network.data.m <- network.data %>%    
  mutate(src.age.cat = cut(src_age_int, c(15,20,25,30,35,40,45,50,55,60,200), right=F)) %>%
  mutate(dest.age.cat = cut(dest_age_int, c(15,20,25,30,35,40,45,50,55,60,200), right=F)) %>%
  mutate(src.age.cat3 = cut(src_age_int, c(15,25,35,200), right=F)) %>%
  mutate(dest.age.cat3 = cut(dest_age_int, c(15,25,35,200), right=F)) %>%
  mutate(acq.age=src_age_int-(YEAR-acq.year)) %>% #age of source when they acquired HIV
  mutate(TimeToTransmission = YEAR - acq.year) %>%
  mutate(MonthsToTransmission = ceiling((YEAR - acq.year)*12)) %>% #month within which transmission occurred (e.g., 1 = transmission ocrurred within < 1 month of acq)
  mutate(YearsToTransmission = ceiling((YEAR - acq.year))) %>% #year within which transmission occurred (e.g., 1 = transmission occurred within < 1 year of acq)
  mutate(era = case_when(Year2 < 2016 ~ "Pre-UTT", TRUE ~ "UTT era")) %>% mutate(Year= Year2) %>%
  mutate(src.stage.f = factor(SRC_STAGE, levels = c(1,2,3,4), labels = c("Acute","Latent","AIDS","on ART"))) %>%
  dplyr::select(-c(SRC_STAGE, Year2)) 
network.data.m$scenario = factor(network.data.m$scenario, levels=c('baseline','ART100pct'))
table(network.data.m$scenario)

labs <- c("1" = "Women", "0" = "Men")
labs.age3 <- c("[15,25)"="15-24","[25,35)"="25-34","[35,200)"="35+")
```
## Model variables

#Transmission variables
```{r, echo = FALSE}
var_guide_baseline <- data.frame(
  var_name = c(names(network.data.m)),
  descr = c(
    "Source transmission ID",
    "Source sex (0 = Male, 1 = Female)",
    "Destination transmission ID",
    "Destiation sex (0 = Male, 1 = Female)",
    "Age of source (years)",
    "Age of destination (years)",
    "Year in sim time-steps",
    "Simulation ID (N=250)",
    "Model scenario (baseline=status quo ART, ART100pct=All HIV+ initiate ART within ~ 6 mo, viral load suppression within 1 year of acquisition",
    "Year transmitter (dest) acquired HIV (NA if seeded in)",
    "5-yr age category of source",
    "5-yr age category of destination",
    "10-yr age category of source",
    "10-yr age category of destination",
    "Age when transmitter acquired HIV (NA if seeded in)",
    "Time (years with decimal) from HIV acquisition to transmission (from source persepctive)",
    "Time (months) from HIV acquisition to transmission (from source persepctive)",
    "Time (years) within which HIV acquisition to transmission occured (from source persepctive)",
    "era (pre/post UTT)",
    "Year transmission occured",
    "Source HIV stage (Acute = untreated acute stage (2.9 months), Latent=untreated latent HIV (>3 months), AIDS=untreated AIDS, on ART=On ART"
    )
  )

network.vars <-kable(
  var_guide_baseline,
  col.names = c("Variable name",
    "Description")
) %>% 
  kable_styling()
network.vars
```

#ReportHIVbyageandgender
#Brings in reporthivbyageandgender.csv. This file defines the population at risk, new infections, and prevalent infections at each time step by strata.

```{r, echo = FALSE}
reporthivbyageandgender <- reporthivbyageandgender %>% 
  mutate(SRC_GENDER = Gender) %>%
  mutate(Year = Year2) %>%
  mutate(NeverOnART = case_when(IP_Key.ARTstate == "NeverOnART" ~ 1, TRUE ~ 0)) %>%
  mutate(src.age.cat3 = cut(Age, c(15,25,35,200), right=F)) %>%
  mutate(src.age.cat = cut(Age, c(15,20,25,30,35,40,45,50,55,60,200), right=F)) %>%
  dplyr::select(-c(Year2, Gender, Age)) 

var_guide_reporthiv <- data.frame(
  var_name = c(names(reporthivbyageandgender)),
  descr = c(
    "Is Circumcised (1=Yes, 0 =No)",
    "ART status among HIV+ (NeverOnART=Never initiated ART, OffART=Initiated but currently off, OnART=On ART)",
    "Newly infected with HIV (n)",
    "Transmitters (n)",
    "Uninfected (n)",
    "Currently infected with HIV (n)",
    "Total population (N)",
    "Simulation ID (N=250)",
    "Model scenario (baseline=status quo ART, ART100pct=All HIV+ initiate ART within ~ 6 mo, viral load suppression within 1 year of acquisition",
    "Sex (0 = Male, 1 = Female)",
    "Calendar year",
    "Never initiated ART (HIV+)",
    "Age category (5-year bins)",
    "Age category (10-year bins)"
  )
)

reportHIV.vars <- kable(
  var_guide_reporthiv,
  col.names = c("Variable name",
    "Description")
) %>% 
  kable_styling()
```

## Contribution to incidence over time by stage 
## of infection and age at transmission
The composition of prevalent HIV infections (stage of infection) changes over time as the epidemic matures and interventions scale. Initially, most infections were caused by acutely infected individuals and infections from high-risk groups.  Over time, infections accumulated in the latent stages and shifted to lower-risk, community-level transmission.  ART scale-up directly reduced the contribution to incidence from those with untreated latent stage infection and untreated AIDS. Incidence attributed to transmission from acutely infected PLHIV also declined in the ART era primarily from the indirect effects of fewer new infections (and thus a lower prevalence of acutely infected individuals) – not from directly suppressing individuals in the acute stage.  Transmissions from PLHIV in the acute stage accounted for 10% of total incidence (<0.5 infections per 100 py) prior to ART scale-up.  After UTT, incidence attributed to PLHIV in the acute stage declined to 0.25 per 100 py, though made up a larger overall proportion of the total incidence (15%).  

```{r, echo = FALSE, fig.width=10, fig.height=7}

#stage
t1 <- network.data.m  %>%
  group_by(scenario, sim.id, Year, src.stage.f) %>%
  summarize(transmissions=n()) 
r1 <- subset(reporthivbyageandgender)  %>%
  group_by(scenario, sim.id, Year) %>%
  summarize(Uninfected.Population=sum(Uninfected.Population)) 
ttable <- merge(t1, r1, by=c("scenario","Year","sim.id"), all.x=T) %>%
  mutate(inc = transmissions/Uninfected.Population) %>%
  group_by(scenario, Year, sim.id) %>%
  mutate(freq = transmissions / sum(transmissions)) %>%
  group_by(scenario, Year, src.stage.f) %>%
  summarize(prop.med=median(freq), prop.lb = quantile(freq, probs = 0.025), prop.ub = quantile(freq, probs = 0.975), inc.med=median(inc), inc.lb = quantile(inc, probs = 0.025), inc.ub = quantile(inc, probs = 0.975)) 

#age
t1 <- network.data.m  %>%
  group_by(scenario, sim.id, Year,SRC_GENDER, src.age.cat3) %>%
  summarize(transmissions=n()) 
r1 <- subset(reporthivbyageandgender)  %>%
  group_by(scenario, sim.id, Year) %>%
  summarize(Uninfected.Population=sum(Uninfected.Population)) 
ttable.age <- merge(t1, r1, by=c("scenario","Year","sim.id"), all.x=T) %>%
  mutate(inc = transmissions/Uninfected.Population) %>%
  group_by(scenario, Year, sim.id) %>%
  mutate(freq = transmissions / sum(transmissions)) %>%
  group_by(scenario, Year,SRC_GENDER, src.age.cat3) %>%
  summarize(prop.med=median(freq), prop.lb = quantile(freq, probs = 0.025), prop.ub = quantile(freq, probs = 0.975), inc.med=median(inc), inc.lb = quantile(inc, probs = 0.025), inc.ub = quantile(inc, probs = 0.975)) 

#Incidence contribution (absolute) by HIV stage

#Stacked bar plot
ggplot(subset(ttable), aes(fill=src.stage.f, y=inc.med*100, x=Year)) + 
  geom_bar(position="stack", stat="identity") +
  scale_x_continuous(expand = c(0,0), breaks=seq(1980,2045,5)) +
  scale_y_continuous(breaks=seq(0,0.035,0.005)*100,expand = c(0,0),"new infections / 100 py") +
  facet_grid(~scenario) +
  coord_cartesian(xlim=c(1980, 2050)) + ggtitle("HIV incidence contribution by source stage of infection") +
  theme(strip.background = element_rect(colour="white", fill="white"), axis.text.x = element_text(angle = 90),legend.title=element_blank(),axis.line = element_line(colour = "black"),panel.background = element_blank(),legend.position = "bottom") +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+ annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

#Line plot 
ggplot(subset(ttable), aes(x=Year)) + 
  geom_line(aes(y=inc.med*100, color=src.stage.f), size=2) +
  geom_ribbon(aes(ymin=inc.lb*100, ymax=inc.ub*100, fill=src.stage.f), size=2, alpha=0.2) +
  scale_x_continuous(expand = c(0,0), breaks=seq(1980,2045,5)) +
  scale_y_continuous(breaks=seq(0,0.035,0.005)*100,expand = c(0,0),"new infections / 100 py") +
  facet_grid(~scenario) +
  coord_cartesian(xlim=c(1980, 2050)) + ggtitle("HIV incidence contribution by source stage of infection") +
  theme(strip.background = element_rect(colour="white", fill="white"), axis.text.x = element_text(angle = 90),legend.title=element_blank(),axis.line = element_line(colour = "black"),panel.background = element_blank(),legend.position = "bottom") +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+ annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

#showing just 2016 and beyond to distill differences between scenarios
ggplot(subset(ttable, Year>2015), aes(x=Year)) + 
  geom_line(aes(y=inc.med*100, color=src.stage.f,linetype=scenario), size=2) +
  #geom_ribbon(aes(ymin=inc.lb*100, ymax=inc.ub*100, fill=src.stage.f), size=2, alpha=0.2) +
  scale_x_continuous(expand = c(0,0), breaks=seq(1980,2045,5)) +
  scale_y_continuous(breaks=seq(0,0.5,0.1),expand = c(0,0),"new infections / 100 py") +
  #facet_grid(~scenario) +
  coord_cartesian(xlim=c(2016, 2050)) + ggtitle("HIV incidence contribution by source stage of infection") +
  theme(strip.background = element_rect(colour="white", fill="white"), axis.text.x = element_text(angle = 90),legend.title=element_blank(),axis.line = element_line(colour = "black"),panel.background = element_blank(),legend.position = "bottom") +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+ annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

#Incidence contribution (proportion) by HIV stage

#Stacked bar plot
ggplot(subset(ttable), aes(fill=src.stage.f, y=prop.med*100, x=Year)) + 
  geom_bar(position="stack", stat="identity") +
  scale_x_continuous(expand = c(0,0), breaks=seq(1980,2045,5)) +
  scale_y_continuous("Proportion of transmissions") +
  facet_grid(~scenario) +
  coord_cartesian(xlim=c(1980, 2050)) + ggtitle("HIV incidence contribution by source stage of infection") +
  theme(strip.background = element_rect(colour="white", fill="white"), axis.text.x = element_text(angle = 90),legend.title=element_blank(),axis.line = element_line(colour = "black"),panel.background = element_blank(),legend.position = "bottom") +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+ annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

#Age contribution stacked bar plots 

ggplot(subset(ttable.age), aes(fill=interaction(src.age.cat3, SRC_GENDER), y=inc.med*100, x=Year)) + 
  geom_bar(position="stack", stat="identity") +
  scale_x_continuous(expand = c(0,0), breaks=seq(1980,2045,5)) +
  scale_y_continuous(breaks=seq(0,0.035,0.005)*100,expand = c(0,0),"new infections / 100 py") +
  facet_grid(~scenario) +
    scale_fill_manual(values = c(brewer.pal(n=3, name="Blues"), brewer.pal(n=3, name="Reds")),labels = c("Male 15-24", "Male 25-34", "Male 35+", "Female 25-34", "Female 25-34", "Female 35+"))+
  coord_cartesian(xlim=c(1980, 2050)) + ggtitle("HIV incidence contribution by source stage of infection") +
  theme(strip.background = element_rect(colour="white", fill="white"), axis.text.x = element_text(angle = 90),legend.title=element_blank(),axis.line = element_line(colour = "black"),panel.background = element_blank(),legend.position = "bottom") +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+ annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

ggplot(subset(ttable.age), aes(fill=interaction(src.age.cat3, SRC_GENDER), y=prop.med*100, x=Year)) + 
  geom_bar(position="stack", stat="identity") +
  scale_x_continuous(expand = c(0,0), breaks=seq(1980,2045,5)) +
  scale_y_continuous("Proportion of total transmissions") +
  facet_grid(~scenario) +
  scale_fill_manual(values = c(brewer.pal(n=3, name="Blues"), brewer.pal(n=3, name="Reds")),labels = c("Male 15-24", "Male 25-34", "Male 35+", "Female 25-34", "Female 25-34", "Female 35+"))+
  coord_cartesian(xlim=c(1980, 2050)) + ggtitle("Proportion contribution to total incidence by age-sex groups") +
  theme(strip.background = element_rect(colour="white", fill="white"), axis.text.x = element_text(angle = 90),legend.title=element_blank(),axis.line = element_line(colour = "black"),panel.background = element_blank(),legend.position = "bottom") +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+ annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  guides(fill = guide_legend(nrow = 1))

```

## Transmisison flows

```{r, fig.width=8, fig.height=8}
#Plots are stratified on UTT era - chose one and then plot

transmission.table.simple <- subset(network.data.m, era=="UTT era")  %>%
  group_by(SRC_GENDER, DEST_GENDER, src.age.cat3, dest.age.cat3) %>%
  summarize(transmissions=n())

##
transmission.table.simple$source.gender.age <- paste(transmission.table.simple$SRC_GENDER, transmission.table.simple$src.age.cat3)
transmission.table.simple$destination.gender.age <- paste(transmission.table.simple$DEST_GENDER, transmission.table.simple$dest.age.cat3)

flow.data <- data.frame("orig_reg" = transmission.table.simple$source.gender.age, "dest_reg" = transmission.table.simple$destination.gender.age, "Transmissions" = transmission.table.simple$transmissions)
table(flow.data$orig_reg)

flow.data$source <- as.numeric(as.factor(flow.data$orig_reg))-1 #must be zero indexed so subtract 1
flow.data$target <- as.numeric(as.factor(flow.data$dest_reg))-1 #must be zero indexed so subtract 1
head(flow.data)

plot.data = data.frame("source"=flow.data$source, "target"=flow.data$target, "value"=flow.data$Transmissions)
nodes <- data.frame("name"=c("Male 15-24","Male 25-34","Male 35+",
                             "Female 15-24","Female 25-34","Female 35+"))
plot.data.m <- plot.data[plot.data$source<3,] #plot for m -> f
plot.data.f <- plot.data[plot.data$source>2,] #plot for f -> m

sankeyNetwork(Links = plot.data.m, Nodes = nodes,
              Source = "source", Target = "target",
              Value = "value", NodeID = "name", nodeWidth = 30, fontSize = 24)

sankeyNetwork(Links = plot.data.f, Nodes = nodes,
              Source = "source", Target = "target",
              Value = "value", NodeID = "name", nodeWidth = 30, fontSize = 24)

```

## Number of transmissions per infected per year

```{r, fig.width=12, fig.height=8}
#All ages
t1 <- network.data.m  %>%
  group_by(scenario, sim.id, SRC_GENDER, Year) %>%
  summarize(transmissions=n()) 
r1 <- subset(reporthivbyageandgender, Year>1981)  %>%
  group_by(scenario, sim.id, SRC_GENDER, Year) %>%
  summarize(Infected=sum(Infected)) 
ttable <- merge(t1, r1, by=c("scenario","SRC_GENDER","Year","sim.id"), all.x=T) %>%
  mutate(inc = transmissions/Infected) %>% #incidence of transmission among infected
  group_by(scenario, SRC_GENDER, Year) %>%
  summarize(inc.med=median(inc), inc.lb = quantile(inc, probs = 0.025), inc.ub = quantile(inc, probs = 0.975)) 

labs <- c("1" = "Women", "0" = "Men")
ggplot(subset(ttable)) + 
  geom_line(aes(y=inc.med, x=Year, color=factor(SRC_GENDER), group=interaction(factor(SRC_GENDER),scenario),linetype=scenario), size=2) +
  geom_ribbon(aes(ymin=inc.lb, ymax=inc.ub, x=Year, fill=factor(SRC_GENDER), group=interaction(factor(SRC_GENDER),scenario),alpha=scenario),alpha=0.2) +
  #facet_grid(~scenario)+
  scale_x_continuous(expand = c(0,0), breaks=seq(1980,2050,5)) +
  scale_y_continuous(expand = c(0,0), breaks=seq(0,0.6,0.05)) + 
  coord_cartesian(xlim=c(1982, 2050), ylim=c(0, 0.5)) +
  xlab("Year")+ ylab("Transmissions per infected individual per year") +
  scale_color_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
  scale_fill_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
  ggtitle("Risk of secondary transmission by sex") +
  theme(strip.background = element_rect(colour="black", fill="white")) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.background = element_blank(),
        legend.position = "bottom")

#By age-group
t1 <- network.data.m  %>%
  group_by(scenario, sim.id,  SRC_GENDER, src.age.cat3, Year) %>%
  summarize(transmissions=n()) 
r1 <- subset(reporthivbyageandgender, Year>1981)  %>%
  group_by(scenario, sim.id, SRC_GENDER, src.age.cat3, Year) %>%
  summarize(Infected=sum(Infected)) 
ttable <- merge(t1, r1, by=c("scenario","SRC_GENDER","src.age.cat3", "Year","sim.id"), all.x=T) %>%
  mutate(inc = transmissions/Infected) %>% #incidence of transmission among infected
  group_by(scenario, SRC_GENDER, src.age.cat3, Year) %>%
  summarize(inc.med=median(inc), inc.lb = quantile(inc, probs = 0.025), inc.ub = quantile(inc, probs = 0.975)) 

labs <- c("1" = "Women", "0" = "Men")
ggplot(subset(ttable)) + 
  geom_line(aes(y=inc.med, x=Year, color=factor(SRC_GENDER), group=interaction(factor(SRC_GENDER),scenario), linetype=scenario), size=2) +
  #geom_ribbon(aes(ymin=inc.lb, ymax=inc.ub, x=Year, fill=factor(SRC_GENDER), group=interaction(factor(SRC_GENDER),scenario)),alpha=0.3) +
  facet_grid(~src.age.cat3, labeller = labeller(src.age.cat3=labs.age3))+
  scale_x_continuous(expand = c(0,0), breaks=seq(1980,2050,5)) +
  scale_y_continuous(expand = c(0,0), breaks=seq(0,0.6,0.05)) + 
  coord_cartesian(xlim=c(1982, 2050), ylim=c(0, 0.3)) +
  xlab("Year")+ ylab("Transmissions per infected individual per year") +
  scale_color_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
  scale_fill_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
  ggtitle("Risk of secondary transmission by sex") +
  theme(strip.background = element_rect(colour="black", fill="white")) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        legend.title=element_blank(), axis.text.x = element_text(angle = 90),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.background = element_blank(),
        legend.position = "bottom")
```

## Distribution of transmission events by time (years) from HIV acquisition

```{r, fig.width=15, fig.height=8}
network.vars
network.data.m$TimeToTransmission <- network.data.m$YEAR - network.data.m$acq.year #time from HIV acquisition until transmission event
hist(network.data.m$TimeToTransmission[network.data.m$SRC_GENDER==0])
hist(network.data.m$TimeToTransmission[network.data.m$SRC_GENDER==1])

#Time to transmission density (among all transmission)
class(network.data.m$TimeToTransmission)
d2 <- subset(network.data.m, is.na(TimeToTransmission)==F) %>%
  group_by(scenario, era, SRC_GENDER, TimeToTransmission) %>%
  summarize(median = quantile(TimeToTransmission, probs = 0.5))

labs <- c("1" = "Women", "0" = "Men")
ggplot(subset(network.data.m, scenario=="baseline"), aes(x=TimeToTransmission, fill=factor(SRC_GENDER), color=factor(SRC_GENDER))) +
  geom_density(alpha=0.3, adjust=3) +
  facet_grid(era~SRC_GENDER, labeller=labeller(SRC_GENDER = labs)) +
  #scale_color_manual(values = c("blue3","red2"), labels=c("Men","Women")) +
  scale_fill_manual(values = c("blue3","red2"), labels=c("Men","Women")) +
  ggtitle("Distribution of time to transmission") +
  theme_bw(base_size=24) +  xlab("Years")+ 
  scale_x_continuous(expand = c(0,0),breaks=seq(0,35,5)) + scale_y_continuous(expand = c(0,0)) + coord_cartesian(xlim=c(0, 40)) +
  geom_vline(data = subset(d2, scenario=="baseline"), aes(xintercept = median, color=factor(SRC_GENDER))) +
  theme(panel.border = element_blank(), strip.background = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"), panel.background = element_blank(),legend.position = "bottom", legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

summary(network.data.m$TimeToTransmission) #median 5.2 years to transmission (1.2-11.6 IQR)
summary(network.data.m$TimeToTransmission[network.data.m$SRC_GENDER==0]) #Men: median 4.7 years to transmission (1.0-10.3 IQR)
summary(network.data.m$TimeToTransmission[network.data.m$SRC_GENDER==1]) #Women: median 6 years to transmission (1.8-13.6 IQR)

#Proportion of HIV transmissions occuring soon after acquisition 1,2,3...50 years (transmission perspective)?
network.data.m$TimeToTransmissionCat <- cut(network.data.m$TimeToTransmission, c(0,1,5,10,50,100), right=F, dig.lab=4)
table(network.data.m$TimeToTransmissionCat)
table(network.data.m$era)

#proportion of HIV acquisitions from transmitters with < 1 year from infection
TimeToTransmissionProportions <- subset(network.data.m,  is.na(TimeToTransmissionCat)==F) %>%
  group_by(scenario, sim.id, Year, dest.age.cat3, DEST_GENDER, TimeToTransmissionCat) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  group_by(scenario, Year, dest.age.cat3, DEST_GENDER, TimeToTransmissionCat) %>%
  summarize(median = quantile(freq, probs = 0.5),
            lb = quantile(freq, probs = 0.025),
            ub = quantile(freq, probs = 0.975))
head(TimeToTransmissionProportions)

ggplot(subset(TimeToTransmissionProportions, TimeToTransmissionCat=="[0,1)" & scenario=="baseline")) + 
  geom_point(aes(y=median, x=Year, color=factor(DEST_GENDER), group=factor(DEST_GENDER)), size=2) +
  geom_line(aes(y=median, x=Year, color=factor(DEST_GENDER)), size=2) +
  #geom_errorbar(aes(x=Year, ymin = lb, ymax = ub, color=factor(DEST_GENDER)), width = 0.2) +
  geom_ribbon(aes(ymin=lb, ymax=ub, x=Year, fill=factor(DEST_GENDER), group=factor(DEST_GENDER)),alpha=0.3) +
  facet_grid(~dest.age.cat3, labeller = labeller(dest.age.cat3=labs.age3))+
  scale_x_continuous(expand = c(0,0), breaks=seq(1980,2050,5)) +
  scale_y_continuous(expand = c(0,0), breaks=seq(0,1,0.05)) + 
  coord_cartesian(ylim=c(0, 1)) +
  xlab("Year")+ ylab("Proportion") +
  scale_color_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
  scale_fill_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
  ggtitle("Proportion of HIV acquisitions from transitters infected for < 1 year") +
  theme(strip.background = element_rect(colour="black", fill="white")) +
  theme(panel.border = element_blank(),
        legend.title=element_blank(),
        panel.grid.minor = element_blank(), axis.text.x = element_text(angle = 90),
        axis.line = element_line(colour = "black"),
        panel.background = element_blank(),
        legend.position = "bottom")

#proportion of HIV transmissions from transmitters with < 1 year from infection
TimeToTransmissionProportions <- subset(network.data.m,  is.na(TimeToTransmissionCat)==F) %>%
  group_by(scenario, sim.id, Year, src.age.cat3, SRC_GENDER, TimeToTransmissionCat) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  group_by(scenario, Year, src.age.cat3, SRC_GENDER, TimeToTransmissionCat) %>%
  summarize(median = quantile(freq, probs = 0.5),
            lb = quantile(freq, probs = 0.025),
            ub = quantile(freq, probs = 0.975))
head(TimeToTransmissionProportions)

ggplot(subset(TimeToTransmissionProportions, TimeToTransmissionCat=="[0,1)" & scenario=="baseline")) + 
  geom_point(aes(y=median, x=Year, color=factor(SRC_GENDER), group=factor(SRC_GENDER)), size=2) +
  geom_line(aes(y=median, x=Year, color=factor(SRC_GENDER)), size=2) +
  #geom_errorbar(aes(x=Year, ymin = lb, ymax = ub, color=factor(SRC_GENDER)), width = 0.2) +
  geom_ribbon(aes(ymin=lb, ymax=ub, x=Year, fill=factor(SRC_GENDER), group=factor(SRC_GENDER)),alpha=0.3) +
  facet_grid(~src.age.cat3, labeller = labeller(src.age.cat3=labs.age3))+
  scale_x_continuous(expand = c(0,0), breaks=seq(1980,2050,5)) +
  scale_y_continuous(expand = c(0,0), breaks=seq(0,1,0.05)) + 
  coord_cartesian(ylim=c(0, 1)) +  theme_bw(base_size=20) +
  xlab("Year")+ ylab("Proportion") +
  scale_color_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
  scale_fill_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
  ggtitle("Proportion of HIV transmissions from transmitters infected for < 1 year") +
  theme(strip.background = element_rect(colour="black", fill="white")) +
  theme(panel.border = element_blank(),
        legend.title=element_blank(),
        panel.grid.minor = element_blank(), axis.text.x = element_text(angle = 90),
        axis.line = element_line(colour = "black"),
        panel.background = element_blank(),
        legend.position = "bottom")

```

## Estimate Reff within the first year of acquisition (How many secondary transmissions occur among those infected for < 1 year?)

```{r, fig.width=18, fig.height=12}

t1 <- network.data.m %>% 
  filter(!is.na(acq.year)) %>% mutate(acq.year = floor(acq.year)) %>%
  group_by(scenario, sim.id, SRC_GENDER, acq.year, MonthsToTransmission) %>%
  summarize(transmissions = n()) 
t1.all <- network.data.m %>% #now aggregate across both genders
  filter(!is.na(acq.year)) %>% mutate(acq.year = floor(acq.year)) %>%
  group_by(scenario, sim.id, acq.year, MonthsToTransmission) %>%
  summarize(transmissions = n()) 
date.of.acquisition2 <- data.frame("SRC_ID"=transmissionreport$DEST_ID, "SRC_GENDER"=transmissionreport$DEST_GENDER, "acq.year"=transmissionreport$YEAR,  "sim.id"=transmissionreport$sim.id, "scenario"=transmissionreport$scenario) #source id used to merge in with source of transmitters
date.of.acquisition2.all <- data.frame("SRC_ID"=transmissionreport$DEST_ID, "acq.year"=transmissionreport$YEAR,  "sim.id"=transmissionreport$sim.id, "scenario"=transmissionreport$scenario) #source id used to merge in with source of transmitters, this is aggregated across both genders
r1 <- date.of.acquisition2  %>% mutate(acq.year = floor(acq.year)) %>%
  group_by(scenario, sim.id, SRC_GENDER, acq.year) %>% summarize(count=n()) 
r1.all <- date.of.acquisition2.all  %>% mutate(acq.year = floor(acq.year)) %>% #do for both genders
  group_by(scenario, sim.id, acq.year) %>% summarize(count=n()) 
ttable <- left_join(t1,r1,by=c("scenario","sim.id","SRC_GENDER","acq.year"))
ttable.all <- left_join(t1.all,r1.all,by=c("scenario","sim.id","acq.year"))
ttable <- ttable %>% mutate(transperperson = transmissions/count) 
ttable.all <- ttable.all %>% mutate(transperperson = transmissions/count, SRC_GENDER=2)  
ttable.merge <- rbind(ttable,ttable.all)

timeframe = 30*12 #time horizon in months to calculate Reff
#fill in missing month and add zeros if there were not transmission events during that month
ttable.selection1 <- complete(subset(ttable.merge, MonthsToTransmission<timeframe & acq.year%in%c(2020)), 
                             MonthsToTransmission = 1:timeframe, fill = list(transperperson = 0))
ttable.selection <- ttable.selection1  %>%
  group_by(sim.id, SRC_GENDER, acq.year) %>% #get cumulative sum
  mutate(reff = cumsum(transperperson), scenario_f = factor(scenario, levels=c('baseline',"ART100pct"))) %>%
  group_by(sim.id, SRC_GENDER, acq.year) %>%
  gather(Metric, Value, transperperson:reff)

ttable.selection.med <- ttable.selection %>% group_by(scenario, SRC_GENDER, acq.year, MonthsToTransmission, Metric) %>% summarize(med.Value=mean(Value))

#Cumulative Reff over time (years)
labs <- c("2"="All", "1" = "Women", "0" = "Men")
labs.metric <- c("transperperson"="Year-specific R_effective", "reff" = "cumulative R_effective")
labs.scenario <- c("baseline"="90-90-90", "ART100pct" = "100% ART init.")
ggplot(subset(ttable.selection, Metric=="reff")) + 
  geom_point(aes(y=Value, x=MonthsToTransmission/12, color=factor(SRC_GENDER), group=interaction(sim.id, SRC_GENDER)), size=1,  alpha=0.01) +
  geom_line(aes(y=Value, x=MonthsToTransmission/12, color=factor(SRC_GENDER), group=interaction(sim.id, SRC_GENDER)), size=1, alpha=0.01) +
  geom_smooth(aes(x=MonthsToTransmission/12, y=Value, group=factor(SRC_GENDER), color=factor(SRC_GENDER)),method="loess", span=0.05, se = F, size=2, linetype=1) +
  facet_wrap(~scenario_f, labeller=labeller(scenario_f=labs.scenario))+
  geom_hline(yintercept=1, linetype=2) +
  scale_x_continuous(expand = c(0,0), breaks=seq(1,timeframe/12,1)) + 
  #scale_y_continuous(expand = c(0,0), breaks=seq(0,1.5,0.1)) + coord_cartesian(ylim=c(0, 1.5)) +
  xlab("Years since HIV acquisition")+ ylab("Reff") +
  scale_color_manual(values = c("blue", "red","purple"), labels = c("Male", "Female","All"))+
  #scale_fill_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
  ggtitle("Number of secondary transmissions per infected individual by year from HIV acquisition") +
  theme(strip.background = element_rect(colour="black", fill="white")) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.background = element_blank(),
        legend.position = "bottom") +  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

#Reff by month and cumulative Reff over time (months)
ggplot(subset(ttable.selection, MonthsToTransmission<25 & scenario=="baseline")) + 
  geom_point(aes(y=Value, x=MonthsToTransmission, color=factor(SRC_GENDER), group=interaction(sim.id, SRC_GENDER)), size=1,  alpha=0.03) +
  geom_line(aes(y=Value, x=MonthsToTransmission, color=factor(SRC_GENDER), group=interaction(sim.id, SRC_GENDER)), size=1, alpha=0.03) +
  geom_line(data=subset(ttable.selection.med,MonthsToTransmission<25& scenario=="baseline"), aes(y=med.Value, x=MonthsToTransmission, color=factor(SRC_GENDER), group=interaction(SRC_GENDER)), size=1) +
  #geom_smooth(aes(x=MonthsToTransmission, y=Value, group=factor(SRC_GENDER), color=factor(SRC_GENDER)),method="loess", span=0.2, se = F, size=2, linetype=1) +
  facet_wrap(~Metric, scales="free", labeller=labeller(Metric=labs.metric, scenario_f=labs.scenario))+
  #geom_hline(yintercept=1) +
  scale_x_continuous(expand = c(0,0), breaks=seq(1,timeframe,1)) + 
  scale_y_continuous(expand = c(0,0)) +
  xlab("Years since HIV acquisition")+ ylab("Reff") +
  scale_color_manual(values = c("blue", "red","purple"), labels = c("Male", "Female","All"))+
  #scale_fill_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
  ggtitle("Number of transmissions per infected individual by month from HIV acquisition") +
  theme(strip.background = element_rect(colour="black", fill="white")) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        legend.title=element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.background = element_blank(),
        legend.position = "bottom") +  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

# ggplot(subset(ttable.selection, Metric=="reff")) + 
#   geom_point(aes(y=Value, x=acq.year, color=factor(MonthsToTransmission), group=interaction(sim.id, acq.year)), size=1,  alpha=0.05) +
#   geom_line(aes(y=Value, x=acq.year, color=factor(MonthsToTransmission), group=interaction(sim.id, acq.year)), size=1, alpha=0.05) +
#   #geom_smooth(aes(x=MonthsToTransmission, y=reff),method="loess", span=0.4, se = F, size=1, linetype=1, color="black") +
#   facet_grid(scenario~SRC_GENDER, labeller=labeller(SRC_GENDER=labs))+
#   scale_x_continuous(breaks=seq(0,timeframe,1)) + 
#   scale_y_continuous(expand = c(0,0), breaks=seq(0,1,0.05)) + 
#   coord_cartesian(ylim=c(0, 1)) +
#   xlab("Months since HIV acquisition")+ ylab("Transmissions per infected individual") +
#   #scale_color_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
#   #scale_fill_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
#   ggtitle("Number of transmissions per infected individual by month from HIV acquisition") +
#   theme(strip.background = element_rect(colour="black", fill="white")) +
#   theme(panel.border = element_blank(), panel.grid.major = element_blank(),
#         legend.title=element_blank(),
#         panel.grid.minor = element_blank(),
#         axis.line = element_line(colour = "black"),
#         panel.background = element_blank(),
#         legend.position = "bottom") +  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
#   annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)
# 

#iterate over all year/time horizon combinations
t1 <- network.data.m %>% 
  filter(!is.na(acq.year)) %>% mutate(acq.year = floor(acq.year)) %>%
  group_by(scenario, sim.id, SRC_GENDER, acq.year, YearsToTransmission) %>%
  summarize(transmissions = n()) 
t1.all <- network.data.m %>% #now aggregate across both genders
  filter(!is.na(acq.year)) %>% mutate(acq.year = floor(acq.year)) %>%
  group_by(scenario, sim.id, acq.year, YearsToTransmission) %>%
  summarize(transmissions = n()) 
date.of.acquisition2 <- data.frame("SRC_ID"=transmissionreport$DEST_ID, "SRC_GENDER"=transmissionreport$DEST_GENDER, "acq.year"=transmissionreport$YEAR,  "sim.id"=transmissionreport$sim.id, "scenario"=transmissionreport$scenario) #source id used to merge in with source of transmitters
date.of.acquisition2.all <- data.frame("SRC_ID"=transmissionreport$DEST_ID, "acq.year"=transmissionreport$YEAR,  "sim.id"=transmissionreport$sim.id, "scenario"=transmissionreport$scenario) #source id used to merge in with source of transmitters, this is aggregated across both genders
r1 <- date.of.acquisition2  %>% mutate(acq.year = floor(acq.year)) %>%
  group_by(scenario, sim.id, SRC_GENDER, acq.year) %>% summarize(count=n()) 
r1.all <- date.of.acquisition2.all  %>% mutate(acq.year = floor(acq.year)) %>% #do for both genders
  group_by(scenario, sim.id, acq.year) %>% summarize(count=n()) 
ttable <- left_join(t1,r1,by=c("scenario","sim.id","SRC_GENDER","acq.year"))
ttable.all <- left_join(t1.all,r1.all,by=c("scenario","sim.id","acq.year"))
ttable <- ttable %>% mutate(transperperson = transmissions/count) 
ttable.all <- ttable.all %>% mutate(transperperson = transmissions/count, SRC_GENDER=2)  
ttable.merge <- rbind(ttable,ttable.all)

# 
# t1 <- network.data.m %>% 
#   filter(!is.na(acq.year)) %>% mutate(acq.year = floor(acq.year)) %>%
#   group_by(scenario, sim.id, SRC_GENDER, acq.year, YearsToTransmission) %>%
#   summarize(transmissions = n()) 
# ttable.years <- left_join(t1,r1,by=c("scenario","sim.id","SRC_GENDER","acq.year"))
# ttable.years <- ttable.years %>% mutate(transperperson = transmissions/count)  

timeframe = 30 #time horizon in years to calculate Reff
ttable.years.selection <- complete(subset(ttable.merge, YearsToTransmission<timeframe+1 & acq.year%in%seq(1980,2020,5)), YearsToTransmission = 1:timeframe, fill = list(transperperson = 0)) #fill in missing month and add zeros if there were not transmission events during that month
ttable.yearscumsum.cont <- 
  ttable.years.selection %>% 
  group_by(scenario, sim.id, SRC_GENDER, acq.year) %>% #get cumulative sum
  mutate(reff = cumsum(transperperson), scenario_f = factor(scenario, levels=c('baseline',"ART100pct"))) %>%
  group_by(scenario_f, SRC_GENDER, acq.year, YearsToTransmission) %>%    
  summarize(med.transperperson=mean(transperperson), med.reff=mean(reff))  #get mean across 250 simulations 
#categorize continuous yeartstotransmission into fewer categories for better plotting
ttable.yearscumsum.cont$med.transperperson.cat <- cut(ttable.yearscumsum.cont$YearsToTransmission, c(0,1,2,3,4,5,10,15,20,25,30), right=T)
ttable.yearscumsum <-  ttable.yearscumsum.cont %>% 
  group_by(scenario_f, SRC_GENDER, acq.year, med.transperperson.cat) %>% 
  summarise(med.transperperson =sum(med.transperperson), med.reff = max(med.reff), YearsToTransmission=max(YearsToTransmission))
ttable.yearscumsum$TransmissionYear <- ttable.yearscumsum$YearsToTransmission + ttable.yearscumsum$acq.year

#Reff by year from acquisition and acquisition year
labs <- c("2"="All", "1" = "Women", "0" = "Men")
ggplot(subset(ttable.yearscumsum)) + 
  geom_area(aes(y=med.transperperson, x=acq.year, fill=factor(YearsToTransmission)), size=1, position="stack") +
  #geom_line(aes(y=med.reff, x=acq.year, color=factor(YearsToTransmission)), size=1) +
  facet_grid(scenario_f~SRC_GENDER, labeller=labeller(SRC_GENDER=labs, scenario_f=labs.scenario))+
  #scale_x_continuous(breaks=seq(0,timeframe,1)) + scale_y_continuous(expand = c(0,0), breaks=seq(0,0.30,0.05)) + 
  #coord_cartesian(ylim=c(0, 0.29)) +
  xlab("Year of HIV acquisition")+ ylab("transmissions per infected") +
  geom_hline(yintercept=1, linetype=2) +   theme(text=element_text(size=24))+
  #scale_color_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
  #scale_fill_manual(values = c("blue", "red"), labels = c("Male", "Female"))+
  ggtitle("Number of transmissions per infected individual by year of HIV acquisition and time from acquisition") +
  theme(strip.background = element_rect(colour="black", fill="white")) +
  guides(fill = guide_legend(reverse=TRUE, nrow=1)) + labs(fill = "Years from HIV acquisition") +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),axis.line = element_line(colour = "black"),
        panel.background = element_blank(), legend.position = "bottom") +  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

brks=c(0.5,1,1.5,2)
library(metR)
ggplot(ttable.yearscumsum.cont) + 
  geom_raster(aes(y=YearsToTransmission, x=acq.year, fill=med.reff))+ 
  scale_fill_gradientn(name="Reff",
                       colours = terrain.colors(20)) +
  facet_grid(SRC_GENDER~scenario_f, labeller=labeller(SRC_GENDER=labs, scenario_f=labs.scenario))+
  stat_contour(aes(y=YearsToTransmission, x=acq.year, z=med.reff), color="black", size=1, linetype=1, breaks=brks) +
  geom_text_contour(aes(y=YearsToTransmission, x=acq.year, z=med.reff), size=7, rotate=F, breaks=brks) +
  theme(legend.key.height = unit(2, "cm")) +
  labs(fill = "Reff") +
  scale_x_continuous(expand = c(0,0), breaks=seq(1985,2020,5),"HIV infected cohort") +
  scale_y_continuous(breaks=seq(0,30,5),expand = c(0,0),"Years to transmissions") +
  coord_cartesian(ylim=c(0, 30)) +
  theme(text=element_text(size=24))+
  theme(strip.background = element_rect(colour="white", fill="white")) +
  theme(axis.text.x = element_text(angle = 90)) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)+
  theme(axis.line = element_line(colour = "black"),panel.background = element_blank())

```













